---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# roarutility

<!-- badges: start -->
<!-- badges: end -->

The goal of roarutility is to process ROAR data. It provides convenience functions to make some common cleaning and processing tasks with ROAR data a bit easier.

## Installation

You can install the development version of roarutility from [GitHub](https://github.com/) with:

```{r install, eval-FALSE, output=FALSE}
# install.packages("pak")
pak::pak("kellywentz/roarutility")
```

## Usage 

### roar.read.csv

This is a basic example which shows you how to read in ROAR data and remove opt-outs in one line of code. 

```{r example-roar-read-csv, eval=FALSE}
library(roarutility)
new_data <- roar.read.csv("all_runs.csv", 
              "~/Documents",
              "https://drive.google.com/file/d/11gYLqU5xT-NMDxWXGQj8WfZ8AVA_lFT9/view?usp=drive_link")
```

Notice how the output dataframe has removed all possible opt-outs from the most up-to-date opt-out CSV.  

### clean_strings 

This is a basic example which shows you how clean_strings() takes in data and outputs data that has removed extra characters from assigning organization variables and converted empty strings to NA values.

```{r example-clean-strings, eval=FALSE}
library(roaryutility)
test_df <- data.frame(
  assigning_schools = c("[irNgj3c]", "irNgj3c"),
  age = c("", "6.7")
)

clean_df <- clean_strings(test_df) 
clean_df 
#   assigning_schools  age
# 1           irNgj3c <NA>
# 2           irNgj3c  6.7
```

Notice how the output dataframe has removed the "[]" characters from the assigning_schools variable and has converted the empty string value in age to an NA value. 

### remove_empty_cols 

This is a basic example which shows you how remove_empty_cols() takes in a dataframe with a column with all NA values and outputs data that has removed the columns with all NA values. 

```{r example-remove-empty-cols, eval=FALSE}
library(roarutility)
test_df <- data.frame(
  firstname = c("Jane", "John", NA, "Kelly"),
  lastname = c("Doe", NA, NA, "Smith"),
  middlename = c(NA, NA, NA, NA)
)

clean_df <- remove_empty_cols(test_df) 
names(clean_df) 
# [1] "firstname" "lastname" 
```

Notice how the output dataframe has removed the column "middlename" because it consisted of all NA values. 

### remove_duplicates

This is a basic example which shows you how remove_duplicates() removes all identical rows across every column. 

```{r example-remove-duplicates, eval=FALSE}
library(roarutility) 
test_df <- data.frame(
  assessment_pid = c("123", "456", NA, "789", "123"),
  roarScore = c(45, 32, 34, 10, 45)
)

clean_df <- remove_duplicates(test_df)
clean_df$assessment_pid
#[1] "123" "456" NA    "789"
```

Notice how the last row test_df[5,] was removed because it had identical values across both columns (assessment_pid and roarScore) as the first row test_df[1,]. 

### remove_accounts

Removes all indicated accounts from the dataframe. Reseearchers can read in data and indicate which or all of the following types of accounts they would like to remove from the dataframe. The function defaults to removing test, demo, pilot, and QA accounts and defaults to not removing NA assessment_pid. The function runs through the organization IDs (i.e., assigning_districts, etc.). It also uses string detection to determine if there are any "test", "pilot", "qa", or "demo" strings within the assessment_pid column. Finally, it runs through to determine if there are test or demo using the variables is_test_data and is_demo_data (if these accounts were chosen to be removed). If selected, the function will also remove assessment_pid = NA. 

