---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# roarutility

<!-- badges: start -->
<!-- badges: end -->

The goal of roarutility is to process ROAR data. It provides convenience functions to make some common cleaning and processing tasks with ROAR data a bit easier.

## Installation

You can install the development version of roarutility from [GitHub](https://github.com/) with:

```{r install}
# install.packages("pak")
pak::pak("kellywentz/roarutility")
```

## Usage 

This is a basic example which shows you how to read in ROAR data and remove opt-outs in one line of code. 

```{r example, eval=FALSE}
library(roarutility)
new_data <- roar.read.csv("all_core_runs_2025-05-27.csv", 
              "~/Documents/GitHub/roar-technical-manual/data",
              "https://drive.google.com/file/d/11gYLqU5xT-NMDxWXGQj8WfZ8AVA_lFT9/view?usp=drive_link")
```

Notice how the output dataframe has removed all possible opt-outs from the most up-to-date opt-out CSV. 

